---
title: "Canopy openess"
author: "Maria Lucia Gonzalez Torres"
date: "2025-07-06"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "figures/README-",
  out.width = '60%', 
  fig.align='center',
  warning = FALSE, 
  message= FALSE
)
```

## House keeping
```{r, eval=FALSE}
# if you want, clear the workspace
rm(list = ls())

# set the working directory
setwd("/Your/working/directory/here")

# load useful library
pacman::p_load(hemispheR, # For canopy fish-eye images processing
               dplyr, # For grammar, data manipulation
               terra, # For geographic (spatial) data manipulation
               magick, # For image manipulation
               exifr # For extracting metadata
               )

image_folder <- "NAME_OF_YOUR_FOLDER_CONTAINING_JPEG_IMAGES"
```

```{r, echo = FALSE}
setwd("/Volumes/OneTouch/UdS/FISHEYE/FISHEYE_GIT/")
pacman::p_load(hemispheR, # For canopy fish-eye images processing
               dplyr, # For grammar, data manipulation
               terra, # For geographic (spatial) data manipulation
               magick, # For image manipulation
               exifr # For extracting metadata
               )
image_folder <- "figures"
```

## Import Image 
```{r, eval=FALSE}
image_file <- file.path(image_folder, "IMAGE1.jpg")
# plot the raster image
terra::plotRGB(terra::rast(image_file))
```

```{r, echo=F, message=F, warning=F}
image_file <- file.path(image_folder, "README_FC2_Dicra.jpg")
terra::plotRGB(terra::rast(image_file))
```



## Select metadata information (optional)
```{r}
xmp_data <- read_exif(image_file) %>%
  select(
    FileName,
    CreateDate,
    LensID,
    Megapixels,
    GPSLatitude,
    GPSLongitude,
    GPSAltitude
  )
glimpse(xmp_data)
```

## import_fisheye   
This function imports and prepares a fisheye image for further analysis (e.g. binarization or canopy structure metrics). It crops and resizes the image based on lens characteristics and user preferences.
```{r}
img<-import_fisheye(filename=image_file,
                    channel = 3,
                    circ.mask=list(xc=1890,yc=1512,rc=1510),
                    circular=TRUE,
                    gamma=2.2,
                    stretch=FALSE,
                    display=TRUE,
                    message=FALSE)
```

*filename*  
Character. The input image filename.  

*channel*    
default = 3\    
The blue channel (3 or "B") has the highest contrast between sky and vegetation. Vegetation reflects less blue light, while the sky appears bright — making it easier to segment. This is the standard choice in hemispherical image analysis (especially in gap fraction estimation).

*circ.mask*   
List. If missing, they are automatically calculated (corresponds to half the lower image side);   
The circular mask parameters (xc,yc,rc) to be applied to the image. The (x,y) coordinate of the center and radius (in pixels).

*circular*      
It allows to specify if the fisheye image is circular (circular=TRUE) or fullframe (circular=FALSE).  

*gamma*  
default = 2.2\   
Gamma correction is a nonlinear operation that adjusts image brightness. Most images (especially JPEGs) are saved with gamma encoding (~2.2) to match how human eyes perceive light (human vision is more sensitive to low light). To accurately analyze light intensity (e.g., canopy openness, gap fraction), you need to undo this compression. Use gamma = 1 when working with raw format (e.g., DNG).

*Stretch*  
default = FALSE\   
Contrast stretching is a **relative, image-specific adjustment**. It linearly rescales pixel intensities so that the darkest and brightest values are spread across the full range (e.g., 0–255). This enhances contrast within each image, but it alters the absolute pixel values, so images no longer reflect true differences in light or exposure between photos. So, if you're comparing metrics like gap fraction, canopy openness, or LAI across multiple images, using stretch = TRUE may distort those differences.

*display*  
default = FALSE\   
if TRUE, it plots the image along with the applied mask and a circle radius.

*message*   
default = TRUE\     
If is set to TRUE, it prints the mask used for importing the image.

## binarize_fisheye ##
This function converts a hemispherical (fisheye) image to a binary image, typically for estimating canopy openness or light penetration. Several thresholding methods can be used.
```{r}
img.bw <- binarize_fisheye(img,
                             method = 'Otsu',
                             zonal = FALSE,
                             manual = NULL,
                             display = TRUE,
                             export = FALSE)
```
*img*     
A single layer fisheye image imported by import_fisheye() 

*method*   
Default = 'Otsu'\   
**Thresholding methods** are techniques used in image binarization, where the goal is to separate foreground from background — by converting a grayscale image into a binary one (black and white). The available methods are described at https://imagej.net/plugins/auto-threshold

*zonal*   
Default = FALSE\    
If TRUE, applies thresholding separately within defined zenith zones (N, W, S, E). Useful for zonal analysis of canopy structure.

*manual*   
Default = NULL\   
Manually set a threshold value (e.g., 0.35). Overrides the method parameter if not NULL.

*display*   
Default = FALSE\   
If is set to TRUE, it plots the classified binary image.

*export*   
Default = FALSE\   
If is set to TRUE, it saves the binary fisheye image as tif file.


## gapfrac_fisheye  ##
This function calculates **gap fraction (the proportion of visible sky)** from a binarized fisheye image. It divides the image into zenith rings and azimuthal segments to estimate light penetration.
```{r}
gapfrac <- gapfrac_fisheye(img.bw,
                            maxVZA = 90,
                            lens = "equidistant",
                            startVZA = 0,
                            endVZA = 90,
                            nrings = 5,
                            nseg = 8,
                            display = TRUE,
                            message = TRUE)
```
*img.bw*  
A single layer binarized fisheye image imported by binarize_fisheye()

*maxVZA*    
DEFAULT = 90    
A zenith angle of 90° corresponds to the horizon — the outer edge of the image.  
  
*lens*     
DEFAULT = "equidistant" 
Lens projection type. Use "equidistant" for most fisheye lenses; other options available by typing *list.lensens*  

*startVZA*   
DEFAULT = 0   
Starting zenith angle (in degrees). Often 0°, the image center.  

*endVZA*   
DEFAULT = 70   
It defines how far from the center (zenith = 0°) toward the edge of the fisheye image you're including in calculations. Setting it to 70° excludes the last 20° of the imagehe because the outermost area often contains lens distortion, vignetting, obstructions.
  
*nrings*   
DEFAULT = 7   
Number of zenith rings (concentric circles) used to divide the image vertically. when you're dividing a fisheye image into zenith rings, you're analyzing concentric circles that represent slices of the sky from overhead (0°) down to the horizon (90°), which helps in estimating how much light gets through at different angles.  
  
*nseg*   
DEFAULT = 8   
Number of azimuthal segments (like pie slices) used to divide the image horizontally - compass directions measured in degrees along the horizon.  
  
*message*   
DEFAULT= FALSE   
If set to TRUE, prints progress and summary messages.
  
*display*   
DEFAULT = FALSE   
If set to TRUE, displays the ring and segment layout on the image.



## canopy_fisheye ##

```{r}
canopy_report <- canopy_fisheye(rdfw = gapfrac) %>%
  rename(PhotoID = id) # to match excel datasheet
glimpse(canopy_report)
```
*rdfw*   
Dataframe. The input dataframe generated from gapfrac_fisheye(), which contains gap fraction for zenith and azimuth bins.

### Site Comparison
**DIFN (Diffuse Non-Interceptance)**
What it measures: Percentage of sky visible through canopy gaps (direct proxy for light penetration).
Range: 0% (fully closed canopy) to 100% (open sky).
Best for: Overall light availability under the canopy.

### Temporal Trends
**LX (Light Transmission Index)**
What it measures: Fraction of potential light reaching the forest floor.
Range: 0 (no light) to 1 (full light).
Best for: Comparing relative light levels between sites/dates. 

### Spatial Light Gradients 
**LXG1/LXG2 (Light Gradient Zones)**
What it measures: Light transmission in specific angular zones (e.g., LXG1 = zenith light, LXG2 = oblique angles).
Best for: Analyzing directional light patterns (e.g., sunflecks vs. diffuse light).




